---
title: "Exploring mediation through major bleeding between DOACs and CV events"
author: "Martijn Schuemie, Zhong Yuan, Elliot Barnathan, Jamie Weaver"
date: "`r Sys.Date()`"
output: 
  word_document: 
    toc: yes
    number_sections: yes
bibliography: Design.bib
---

```{r setup, include=FALSE,warning=FALSE}
library(dplyr)
library(knitr)

source("PrintCohortDefinitions.R")

```

# Introduction

Anticoagulant therapies have undergone significant evolution with the development of direct oral anticoagulants (DOACs), which have been increasingly favored over traditional warfarin due to their promising safety profiles, fewer dietary restrictions, and reduced need for monitoring. While DOACs are associated with a lower risk of major bleeding compared to warfarin, their overall impact on cardiovascular (CV) outcomes remains a subject of intense research. Major bleeding is a critical concern in anticoagulant therapy and can profoundly influence patient outcomes and treatment efficacy. 

This real-world study aims to dissect the complex interplay between anticoagulant type, major bleeding events, and cardiovascular outcomes. This protocol outlines the methodology for an observational study that will estimate the extent to which the effect of DOACs versus warfarin on cardiovascular outcomes is mediated through their differential impact on major bleeding rates. By leveraging real-world data, the study will contribute valuable insights into the comparative effectiveness and safety of these anticoagulants, guiding clinical decision-making and potentially informing future guidelines.

This study will utilize novel advanced statistical techniques to model the mediation effects, controlling for a range of confounding factors that could influence both bleeding risk and cardiovascular outcomes.

# Study design

This study will employ a new-user comparative cohort design, comparing a target cohort (rivaroxaban or the entire class of DOACs) to a comparator cohort (warfarin). 
Both target and cohort are defined as first exposure to drug of interest, requiring 365 days of prior observation. 
Followup starts on the day of treatment initiation, and ends at end of continuous exposure (allowing for 30-day gaps) or occurrence of the outcome, whichever comes first. 
A single mediator - major bleeding - is defined and included in the model. This table lists all the comparisons that will be made:

| Target      | Comparator | Mediator       | Outcome         |
|:------------|:-----------|:---------------|:----------------|
| Rivaroxaban | Warfarin   | Major bleeding | Acute MI        |
| Rivaroxaban | Warfarin   | Major bleeding | Ischemic stroke |
| DOACs       | Warfarin   | Major bleeding | Acute MI        |
| DOACs       | Warfarin   | Major bleeding | Ischemic stroke |

The model is a Cox proportional hazards model, with the mediator as time-varying covariate. 
Cohort definitions of the targets, comparator, mediator and outcomes are provided in Appendix A. 

## Estimands of interest

For each target-comparator-outcome triplet, two Cox modes will be fitted, one with the mediator and one without. From these two models, the following estimands will be reported, each on the hazard ratio scale, including their 95% confidence intervals: [@vanderWeele_2011]

- **Main effect**: The effect of the target on the outcome, relative to the comparator.
- **Direct effect**: The effect of the target on the outcome, relative to the comparator, *not* mediated by the mediator.
- **Indirect effect**: The effect of the target on the outcome, relative to the comparator, mediated by the mediator.

The indirect effect is estimated using the difference method, subtracting the (log) direct effect from the (log) main effect.

## Confounding adjustment

Adjustment for confounding between target and comparator is achieved by large-scale propensity scores (LSPS), [@tian_2018] used for variable ratio PS matching. [@rassen_2012] 

Adjustment for confounding between those with the mediator and those without is achieved by a large-scale mediator risk score (MRS). 
The MRS is fitted as a Poisson regression by using the same set of baseline covariates used in the LSPS, including all demographics, drug exposures, conditions, procedures, etc. observed on or in the year prior to treatment initiation. 
Similar to the LSPS, the MRS model is fitted using L2 regularization, using 10-fold cross-validation to select the optimal hyperparameter by optimizing out-of-sample likelihood.
The MRS is included in the outcome model using a 5-knot bicubic spline.

## Negative controls

A set of 50 negative control outcomes - outcomes believed to be caused by neither the target nor the comparator - has been defined. (See Appendix B). We assume that these controls are negative both for the direct effect and the mediated effect.
The negative control summary estimates will be used to estimate residual systematic error for all estimands of interest.


# Data sources

The DatabaseDiagnostics package was used to select those databases that appear to have the elements needed for the example estimation questions:

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, results="asis"}
dbInfo <- readRDS("DatabaseInfo.rds") %>%
  arrange(cdmSourceName)
output <- c()
for (i in seq_len(nrow(dbInfo))) {
  output[i] <- sprintf("\n## %s\n\n %s\n", dbInfo$cdmSourceName[i], dbInfo$sourceDescription[i])
}
writeLines(output)
```

# References

<div id="refs"></div>

# Appendix A: Cohort Definitions

```{r echo=FALSE, results="asis", warning=FALSE, message=FALSE}
cohortDefinitionSet <- readRDS("CohortDefinitionSet.rds")
```

```{r echo=FALSE, results="asis", warning=FALSE, message=FALSE}
json <- cohortDefinitionSet$json[cohortDefinitionSet$cohortId == 16586]
printCohortDefinitionFromNameAndJson("Target: Rivaroxaban", json)
```

```{r echo=FALSE, results="asis", warning=FALSE, message=FALSE}
json <- cohortDefinitionSet$json[cohortDefinitionSet$cohortId == 16329]
printCohortDefinitionFromNameAndJson("Target: DOACs", json)
```

```{r echo=FALSE, results="asis", warning=FALSE, message=FALSE}
json <- cohortDefinitionSet$json[cohortDefinitionSet$cohortId == 16330]
printCohortDefinitionFromNameAndJson("Comparator: Warfarin", json)
```

```{r echo=FALSE, results="asis", warning=FALSE, message=FALSE}
json <- cohortDefinitionSet$json[cohortDefinitionSet$cohortId == 16484]
printCohortDefinitionFromNameAndJson("Mediator: Major bleeding", json)
```

```{r echo=FALSE, results="asis", warning=FALSE, message=FALSE}
json <- cohortDefinitionSet$json[cohortDefinitionSet$cohortId == 2072]
printCohortDefinitionFromNameAndJson("Outcome: Acute MI", json)
```

```{r echo=FALSE, results="asis", warning=FALSE, message=FALSE}
json <- cohortDefinitionSet$json[cohortDefinitionSet$cohortId == 2087]
printCohortDefinitionFromNameAndJson("Outcome: Ischemic stroke", json)
```

# Appendix B: Negative controls
```{r echo=FALSE, warning=FALSE, message=FALSE}
readr::read_csv(
    file = "NegativeControls.csv",
    show_col_types = FALSE
  ) %>%
  arrange(conceptName) %>%
  select("Concept ID" = "conceptId", Name = "conceptName") %>%
  kable(linesep = "", booktabs = TRUE, longtable = TRUE)
```